---
title: "Handin 3.3"
author: "Gr. 5.217b"
date: "4/11/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Opgave 1
Read the data into R and plot the data. Argue that it is not a good idea to use a general linear model

```{r, include = FALSE}
library("readxl")
DATA_HANDIN <- read_excel("DATA_HANDIN.xlsx")
data <- DATA_HANDIN
data$y <- as.numeric(data$y)
```

```{r}
plot(data)
qqnorm(data$y)
qqline(data$y)
```

Det ses af qq-plottet at dataen ikke følger en normal fordeling og det ville derfor være fordel-agtigt ikke at anvende en generel lineær model.


## Opgave 2
Use the glm function to fit a simple inverse Gaussian regression model to the data with the link function used by R. (Hint: In glm, use family=inverse.gaussian(link="1/mu^2")). Extract the estimated $\beta$ parameters

```{r}
y <- data$y
x <- data$x
mu <- mean(y)

mod = glm(formula=y~x,family=inverse.gaussian(link = "1/mu^2"))
coef = coefficients(mod); coef
```


## Opgave 3
Create confidence intervals for the $\beta$ parameters. Do the confidence intervals agree with the model used for simulating the data?
```{r}
confbeta = confint.default(mod); confbeta
```

Da koefficienterne anvendt til at simulere er $2200$ og $-2$ ligger disse inde i konfidensintervallerne fundet ovenover.


## Opgave 4
Plot the estimated curve $\hat{\mu} = g^{-1}(\hat{\beta_0} + \hat{\beta_1}x)$ on top of the data. Does it fit?
```{r}
plot(data)
curve((coef[1] + coef[2]*x)^(1/(-2)),add=TRUE,col="red") 
```


## Opgave 5
Make a goodness of fit test to check the fit of the model.
```{r}
D = deviance(mod); D
1-pchisq(D,length(x)-length(coef))
```



